Spin canonical divisors of log stacky curves

### 2015-06-27 stacky_write_up/spin_cring.tex
more edits to genus 2

\section{Background}
\label{sec:background}
Look at Spencer-Kodaira \cite{kodaira:complex-manifolds}.
Include the following:
\begin{enumerate}
	\item Notation for Coarse space
	\item Notation for Stacky points with associated stabilizer orders (I.e. as those points for which the divisor can have rational coefficients with denominator equal to the stabilizer order)
	\item Note that this result is independence of base field
	\item background on Gr\"{o}bner bases
	\item Notation $h^0 = \dim_k H^0$.
	\item Why we have these fractions $\frac{e_i - 1}{2e_i}$ and why $e_i$ must be odd.
	\item Include notation for $X$ as coarse space and $L_X$ the corresponding log half canonical divisor.
	\item Notation for grobner bases and orderings (include grlex, grevlex, and block order, as defined in \cite[Definition 8.1.1]{vzb:stacky})
\end{enumerate}

@@

The main result of this subsection is that for a log spin curve (with no stacky points) $(X,\Delta,L)$, the half-canonical ring $R_L$ is generated in degree at most $5$, with relations in degree at most $11$ if $\Delta = 0$ and at most $10$ if $\Delta > 0$. The generation bound is shown in Lemma ~\ref{lem:generation_5} and the relations bound is shown in Lemma ~\ref{prop:relations_11}. In fact, it is fairly immediate from Riemann--Roch and the basepoint-free pencil trick that $R_L$ is generated in degree at most $6$, as is demonstrated in the following
Lemma ~\ref{lem:semicanonical_generation}. However, to reduce the bound from 5 to 6 in the case $g =2$, we will have to work a little in Proposition ~\ref{prop:genus_2_generation_5}. The following proof of Lemma ~\ref{lem:semicanonical_generation} is due to Neves \cite[Proposition III.4]{neves:halfcan}, although he omits the fairly tricky genus 2 case. We produce the proof here for completeness.

\begin{lem}
\label{lem:semicanonical_generation}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2 L \sim K + \Delta$. Then, the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective if $n \geq 5$ or if ($n = 5$ and $\Delta > 0$).
In particular, $R_L$ is generated in degree at most $6$ if $\Delta = 0$ and in degree at most $5$ if $\Delta > 0$.
\end{lem}
\begin{proof}
Therefore, to show there are no new generators in degree at least $7$, it suffices to show that if $n \geq 5$ or if ($n = 5$ and $\Delta > 0$), the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective. Indeed, let $s_1,s_2 \in H^0(2L) \cong H^0(K)$ be two independent sections. 
Since $K$ is basepoint-free, so is the vector subspace $V = ks_1 \oplus k s_2 \subset H^0(K)$. 
Then, by the basepoint-free pencil trick, we obtain an exact sequence
$$\begin{tikzcd}
0 \ar r & H^0((n-2)L) \ar r & V \otimes H^0(nL) \ar{r}{f} & H^0((n+2)L)
\end{tikzcd}$$

We wish to show $f$ is surjective.
Note that $\dim_k \ker f = \dim_k H^0((n-2)L) = (n-3)(g - 1)$ using Riemann--Roch and the assumption if $n \geq 5$ or ($n = 5$ and $\Delta > 0$), as in the case $\Delta > 0$ we have $\deg \Delta \geq 2$.
Additionally, $\dim_k V \otimes H^0(nL) = 2 \cdot (n- 1)(g - 1)$, again using Riemann--Roch.
Therefore, $$\dim_k \im f = 2 \cdot (n- 1)(g - 1) -(n-3)(g - 1) = (n+ 1)(g - 1) = \dim_k H^0((n+2)L),$$
again using Riemann-Roch, implying $f$ is surjective.
\end{proof}

\begin{rem}
\todo{This remark isn't quite right, we're changing the proof}
In \cite[Proposition III.4]{neves:halfcan}, Neves claims to prove that $R_L$ is generated in degree at most 5. However, when the proof is carefully examined, it only shows generation in degree at most 6. For this reason, we only state Lemma ~\ref{lem:semicanonical_generation} as generation in degree at most 6. We next proceed to show this can be reduced to generation in degree 5. The trickiest case is genus 2, which we address next.
\end{rem}

@@

There are two main cases: when $\delta \geq 2$ and when $\delta = 0$
 (note that $\delta$ must be even because $\halfcan_X \in \di X$ with $2\halfcan_X \sim K_X +\Delta$, where $\deg K_X$ is even). \todo{make a remark about this somewhere in the beginning and reference it} In the first case, $\deg \lfloor \halfcan \rfloor
\geq 0$ so the result directly follows from Lemma
~\ref{lem:sat-1}. When $\delta = 0$, the proof is a rather involved induction using Lemma
~\ref{lem:sat-3} and Lemma ~\ref{lem:raise-stacky-order}.
The remainder of this section is dedicated to characterizing the
saturations (Subsection ~\ref{ssec:g_0_saturation}), inductive
base cases (Subsection ~\ref{ssec:g_0_base}), and exceptional
cases (Subsection ~\ref{ssec:g-0-exceptional}) when $\delta = 0$
before proving the main theorem in full for all $\delta$ in
Subsection ~\ref{ssec:g-0-main}.

>>>>

\section{Background}
\label{sec:background}
Look at Spencer-Kodaira \cite{kodaira:complex-manifolds}.
Include the following:
\begin{enumerate}
	\item Notation for Coarse space
	\item Notation for Stacky points with associated stabilizer orders (I.e. as those points for which the divisor can have rational coefficients with denominator equal to the stabilizer order)
	\item Note that this result is independence of base field
	\item background on Gr\"{o}bner bases
	\item Notation $h^0 = \dim_k H^0$.
	\item Why we have these fractions $\frac{e_i - 1}{2e_i}$ and why $e_i$ must be odd.
	\item Include notation for $X$ as coarse space and $L_X$ the corresponding log half canonical divisor.
	\item Notation for grobner bases and orderings (include grlex, grevlex, and block order, as defined in \cite[Definition 8.1.1]{vzb:stacky})
	\item Make sure to standardize the notation $\sx,\sx'$ for stacky curves, $L,L'$ for half canonical divisors, and $X,X'$ for (non stacky) curves.
\end{enumerate}

\begin{rem}
\label{rem:delta-not-1}
Suppose $(\sx,\Delta,L)$ is a log spin curve. Note that $\deg \lfloor L \rfloor$ is even. because $2 \cdot \deg L = \deg \Delta + \deg \lfloor K \rfloor = \deg \Delta + 2(g-1).$ Therefore, $\deg \Delta$ is even. In particular, we shall often use $\deg \Delta \neq 1$.
\end{rem}

@@

The main result of this subsection is that for a log spin curve with no stacky points $(X,\Delta,L)$, the half-canonical ring $R_L$ is generated in degree at most $5$, with relations in degree at most $11$ if $\Delta = 0$ and at most $10$ if $\Delta > 0$. The generation bound is shown in Lemma ~\ref{lem:generation_5} and the relations bound is shown in Lemma ~\ref{prop:relation_11}. In fact, it is fairly immediate from Riemann--Roch and the basepoint-free pencil trick that $R_L$ is generated in degree at most $6$, as is demonstrated in the following
Lemma ~\ref{lem:semicanonical_generation}. However, to reduce the bound from 5 to 6 in the case $g =2$, we will have to work a little in Proposition ~\ref{prop:genus_2_generation_5}.

\begin{rem}
The following proof of Lemma ~\ref{lem:semicanonical_generation} is shown in Neves \cite[Proposition III.4]{neves:halfcan}, although he omits the fairly tricky genus 2 case. We produce the proof here for completeness.
\end{rem}

\begin{lem}
\label{lem:semicanonical_generation}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2 L \sim K + \Delta$. Let $s_1,s_2 \in H^0(X,2L)$ be two independent sections such that the vector subspace $V = ks_1 \oplus k s_2 \subset H^0(K)$ is basepoint free. Then, the map
\begin{align*}
	V \otimes H^0(nL) \rightarrow H^0((n+2)L)
\end{align*}
is surjective if $n \geq 5$ and if {\rm(}$n = 4$ and $\Delta > 0${\rm)}.
In particular, $R_L$ is generated in degree at most $6$ if $\Delta = 0$ and in degree at most $5$ if $\Delta > 0$.
\end{lem}
\begin{proof}
To show there are no new generators in degree at least $7$ if $\Delta = 0$ and in degree at least $6$ if $\Delta > 0$, it suffices to show that if $n \geq 5$ or if ($n = 4$ and $\Delta > 0$), the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective. Indeed, since $V= ks_1 \oplus k s_2 \subset H^0(K)$ is basepoint free, by the basepoint-free pencil trick, (see \cite[Lemma 2.6]{saint-donat:proj} for a proof,) we obtain an exact sequence
$$\begin{tikzcd}
0 \ar r & H^0((n-2)L) \ar r & V \otimes H^0(nL) \ar{r}{f} & H^0((n+2)L)
\end{tikzcd}$$

We wish to show $f$ is surjective.
Note that $\dim_k \ker f = \dim_k H^0((n-2)L) = (n-3)(g - 1)$ using Riemann--Roch and the assumption $n \geq 5$ or ($n = 4$ and $\Delta > 0$). Note that here we are implicitly using the Remark ~\ref{rem:delta-not-1}
Additionally, $\dim_k V \otimes H^0(nL) = 2 \cdot (n- 1)(g - 1)$, again using Riemann--Roch.
Therefore, $$\dim_k \im f = 2 \cdot (n- 1)(g - 1) -(n-3)(g - 1) = (n+ 1)(g - 1) = \dim_k H^0((n+2)L),$$
again using Riemann-Roch, implying $f$ is surjective.
\end{proof}

@@

There are two main cases: when $\delta \geq 2$ and when $\delta = 0$, where we are implicitly appealing to Remark ~\ref{rem:delta-not-1}. In the first case, $\deg \lfloor \halfcan \rfloor
\geq 0$ so the result directly follows from Lemma
~\ref{lem:sat-1}. When $\delta = 0$, the proof is a rather involved induction using Lemma
~\ref{lem:sat-3} and Lemma ~\ref{lem:raise-stacky-order}.
The remainder of this section is dedicated to characterizing the
saturations (Subsection ~\ref{ssec:g_0_saturation}), inductive
base cases (Subsection ~\ref{ssec:g_0_base}), and exceptional
cases (Subsection ~\ref{ssec:g-0-exceptional}) when $\delta = 0$
before proving the main theorem in full for all $\delta$ in
Subsection ~\ref{ssec:g-0-main}.
